Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-01-08 Paper/paper.tex
Update text

\begin{figure*}
  \begin{center}
    % \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small.pdf}
    \includegraphics[height=0.88\textheight,keepaspectratio]{mosaic_small.pdf}
  \end{center}
  \caption{Luminosity function estimates from $z=0.3$ to $2.6$.  The
    symbols show our inferred binned luminosity functions.  In each
    redshift bin, the black curve shows the best-fit double power law
    model, which is represented by the median of the posterior
    probability distribution function.  Yellow curves show a random
    draw from the posterior probability distribution function.  As
    discussed in the text, luminosity bins that show a decrease in the
    luminosity function at the faint end are deemed to be incomplete
    are are removed from our analysis while fitting the double power
    law model.  These bins are represented by open symbols in this
    figure.  Each panel shows the number of quasars in the respective
    redshift bin.  The number in parenthesis indicates the total
    number of quasars in the bin, before excluding incomplete
    bins.}
  \label{fig:mosaic}
\end{figure*}

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth]{evolution_alldata.pdf}
  \end{center}
  \caption{Parameter evolution from individual fits in
    Figures~\ref{fig:mosaic}.  Parameters from BOSS quasars are seen
    to be systematically offset.}
  \label{fig:evoln}
\end{figure*}

@@

\begin{figure*}
  \begin{center}
    \includegraphics[height=0.88\textheight,keepaspectratio]{mosaic_small_global.pdf}
  \end{center}
  \caption{Luminosity function estimates from $z=0.3$ to $2.6$.
    Similar to Figure~\ref{fig:mosaic}, the symbols show our inferred
    binned luminosity functions.  In each redshift bin, the yellow
    curves show the best-fit double power law luminosity function in
    that redshift bin.  The green curves show the posterior
    distribution of the global model.  The global model agrees very
    well with the data in all redshift bins, except those spanned by
    BOSS, as BOSS quasars are excluded from the global analysis.}
  \label{fig:mosaic_global}
\end{figure*}

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth]{evolution_global.pdf}
  \end{center}
  \caption{Luminosity function parameter evolution in the global
    model.  The symbols show the best fit parameters with 68\%
    uncertainties from the double power law fits to the data in
    redshift bins.  These points are identical to those in
    Figure~\ref{fig:evoln}, except that bins with BOSS qsos ($z\sim
    2$--$3$) are not shown.  In each panel, the black solid curve
    shows the best fit global model, and the coloured curves show a
    random sample drawn from the global posterior.}
  \label{fig:evoln_global}
\end{figure*}

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{emissivity_18.pdf}
    \includegraphics[width=0.5\textwidth]{emissivity_20.pdf}
    \end{tabular}
  \end{center}
  \caption{LyC emissivity of AGN assuming 100\% escape fraction.
    Model luminosity functions are integrated down to $M_{1450}=-18$
    in the left panel and $-20$ in the right panel.  For comparison,
    various other published determinations of the 1~Ry emissivity of
    quasars are also shown in both panels of Figure~\ref{fig:e912}.
    The blue curve in these panels shows the AGN 1~Ry emissivity model
    of \citet{2012ApJ...746..125H}, which is very similar to the model
    of \citet{2007ApJ...654..731H}.  The red symbols with error bars
    show the determination from the fits of
    \citet{2015AA...578A..83G}.  The maroon curve is the published fit
    to the 1~Ry emissivity evolution in the fits of
    \citet{2017MNRAS.466.1160M}.  The green curve is the model of
    \citet{2015ApJ...813L...8M}.}
  \label{fig:e912}
\end{figure*}

@@

The 1~Ry emissivity can now be used to calculate the ionizing flux and
photoionization rate contributed by quasars.  The frequency dependence
above 1~Ry is given by Equation~(\ref{eqn:sed}) or
(\ref{eqn:sed_faint}), so that for $\nu > \nu_{912}$
\begin{equation}
  \epsilon_\nu = \epsilon_{912}\left(\frac{\nu}{\nu_{912}}\right)^\alpha,
  \label{eqn:epsilon_freq}
\end{equation}
where $\alpha=-1.70$ (Equation~\ref{eqn:sed}) or $-0.56$
(Equation~\ref{eqn:sed_faint}).  The corresponding flux in the IGM is
written as a solution of the cosmological radiative transfer equation
as \citep{2012ApJ...746..125H}
\begin{multline}
  j(\nu_0, z_0)=\frac{1}{4\pi}\int_{z_0}^\infty dz\frac{dl}{dz}\frac{(1+z_0)^3}{(1+z)^3}\epsilon(\nu,z)\\
  \times\exp{(-\tau_\mathrm{eff}(\nu_0, z_0, z))},
  \label{eqn:flux}
\end{multline}
where
\begin{equation}
  \nu = \nu_0\left(\frac{1+z}{1+z_0}\right).
\end{equation}
Here the effective optical depth is estimated as
\begin{equation}
  \tau_\mathrm{eff}(\nu_0, z_0, z) = \int_{z_0}^z dz^\prime\int_0^\infty dN_\mathrm{HI} f(N_\mathrm{HI}, z^\prime) (1-e^{-\tau_\nu}),
\end{equation}
where $\tau_\nu=\sigma_\nu N_\mathrm{HI}$ and the column density
distribution $f(N_\mathrm{HI}, z)$ is taken to be a power law \citep{2013MNRAS.436.1023B}
\begin{equation}
  f(N_\mathrm{HI}, z) = \frac{A}{N_\mathrm{LL}}\left(\frac{N_\mathrm{HI}}{N_\mathrm{LL}}\right)^{-\beta_N}\left(\frac{1+z}{4.5}\right)^{-\beta_z},
\end{equation}
with $A=0.93$, $\beta_N=1.33$, $\beta_z=1.92$, and
$N_\mathrm{LL}=10^{17.2}$~cm$^{-2}$.  The hydrogen photoionization
rate is then given by
\begin{equation}
  \Gamma_\mathrm{HI}=4\pi\int_{\nu_{912}}^\infty d\nu \frac{j(\nu,z)}{h\nu} \sigma(\nu),
\end{equation}
where $\sigma$ is the photoionization cross-section.  The result is
shown in Figure~\ref{fig:gammapi}, in which as before the solid black
curve shows the hydrogen photoionisate rate in our best-fit global
model, and the yellow curves show a random sample from the global
posterior.  Also, as before, the right panel of
Figure~\ref{fig:gammapi} shows the result when we only account for
quasars brighter than $M_{1450}=-20$, while the left panel shows the
result when the integration limit is $M_{1450}=-18$.  The red points
with error bars in this figure show the measurements of
\citet{2013MNRAS.436.1023B}, derived from the Lyman-$\alpha$ forest,
and the orange points show the measurements by
\citet{2011MNRAS.412.2543C} from quasar proximity zones.  The dashed
green curve shows the hydrogen photoionisation rate evolution in the
model of \citet{2012ApJ...746..125H}, while the solid green curve
shows the contribution of quasars to it.

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{g_18.pdf}
    \includegraphics[width=0.5\textwidth]{g_20.pdf}
    \end{tabular}
  \end{center}
  \caption{AGN contribution to the hydrogen photoionisation rate,
    assuming 100\% escape fraction.  Model luminosity functions are
    integrated down to $M_{1450}=-18$ in the left panel and $-20$ in
    the right panel.  The solid black curve shows the hydrogen
    photoionisate rate in our best-fit global model, and the yellow
    curves show a random sample from the global posterior.  The red
    points with error bars in this figure show the measurements of
    \citet{2013MNRAS.436.1023B}, derived from the Lyman-$\alpha$
    forest, and the orange points show the measurements by
    \citet{2011MNRAS.412.2543C} from quasar proximity zones.  The
    dashed green curve shows the hydrogen photoionisation rate
    evolution in the model of \citet{2012ApJ...746..125H}, while the
    solid green curve shows the contribution of quasars to it. }
  \label{fig:gammapi}
\end{figure*}

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{evolution.pdf}
  \end{center}
  \caption{Parameter evolution from individual fits.  Coloured points
    show results when Giallongo quasars are not included.  Black
    points show results when Giallongo quasars are included.}
\end{figure*}

@@

@@

>>>>

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small.pdf}
    %\includegraphics[height=0.88\textheight,keepaspectratio]{mosaic_small.pdf}
  \end{center}
  \caption{Luminosity function estimates from $z=0.3$ to $2.6$.  The
    symbols show our inferred binned luminosity functions.  In each
    redshift bin, the black curve shows the best-fit double power law
    model, which is represented by the median of the posterior
    probability distribution function.  Yellow curves show a random
    draw from the posterior probability distribution function.  As
    discussed in the text, luminosity bins that show a decrease in the
    luminosity function at the faint end are deemed to be incomplete
    are are removed from our analysis while fitting the double power
    law model.  These bins are represented by open symbols in this
    figure.  Each panel shows the number of quasars in the respective
    redshift bin.  The number in parenthesis indicates the total
    number of quasars in the bin, before excluding incomplete
    bins. \textbf{[Remove redshift bins that were excluded from global
        analysis.]}}
  \label{fig:mosaic}
\end{figure*}

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.7\textwidth]{evol_28sep.pdf}
  \end{center}
  \caption{Parameter evolution from individual fits in
    Figures~\ref{fig:mosaic}.  Parameters from BOSS quasars are seen
    to be systematically offset.}
  \label{fig:evoln}
\end{figure*}

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small_global.pdf}
  \end{center}
  \caption{Luminosity function estimates from $z=0.3$ to $2.6$.
    Similar to Figure~\ref{fig:mosaic}, the symbols show our inferred
    binned luminosity functions.  In each redshift bin, the yellow
    curves show the best-fit double power law luminosity function in
    that redshift bin.  The green curves show the posterior
    distribution of the global model.  The global model agrees very
    well with the data in all redshift bins, except those spanned by
    BOSS, as BOSS quasars are excluded from the global analysis.}
  \label{fig:mosaic_global}
\end{figure*}

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.7\textwidth]{evolution_global_9oct.pdf}
  \end{center}
  \caption{Luminosity function parameter evolution in the global
    model.  The symbols show the best fit parameters with 68\%
    uncertainties from the double power law fits to the data in
    redshift bins.  These points are identical to those in
    Figure~\ref{fig:evoln}, except that bins with BOSS qsos ($z\sim
    2$--$3$) are not shown.  In each panel, the black solid curve
    shows the best fit global model, and the coloured curves show a
    random sample drawn from the global posterior.}
  \label{fig:evoln_global}
\end{figure*}

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{emissivity_18_9oct.pdf}
    \includegraphics[width=0.5\textwidth]{emissivity_20_9oct.pdf}
    \end{tabular}
  \end{center}
  \caption{LyC emissivity of AGN assuming 100\% escape fraction.
    Model luminosity functions are integrated down to $M_{1450}=-18$
    in the left panel and $-20$ in the right panel.  For comparison,
    various other published determinations of the 1~Ry emissivity of
    quasars are also shown in both panels of Figure~\ref{fig:e912}.
    The blue curve in these panels shows the AGN 1~Ry emissivity model
    of \citet{2012ApJ...746..125H}, which is very similar to the model
    of \citet{2007ApJ...654..731H}.  The red symbols with error bars
    show the determination from the fits of
    \citet{2015AA...578A..83G}.  The maroon curve is the published fit
    to the 1~Ry emissivity evolution in the fits of
    \citet{2017MNRAS.466.1160M}.  The green curve is the model of
    \citet{2015ApJ...813L...8M}.}
  \label{fig:e912}
\end{figure*}

@@

The 1~Ry emissivity can now be used to calculate the ionizing flux and
photoionization rate contributed by quasars.  The frequency dependence
above 1~Ry is given by Equation~(\ref{eqn:sed}) or
(\ref{eqn:sed_faint}), so that for $\nu > \nu_{912}$
\begin{equation}
  \epsilon_\nu = \epsilon_{912}\left(\frac{\nu}{\nu_{912}}\right)^\alpha,
  \label{eqn:epsilon_freq}
\end{equation}
where $\alpha=-1.70$ (Equation~\ref{eqn:sed}) or $-0.56$
(Equation~\ref{eqn:sed_faint}).  The flux in the IGM is
written as a solution of the cosmological radiative transfer equation
as \citep{2012ApJ...746..125H}
\begin{multline}
  j(\nu_0, z_0)=\frac{1}{4\pi}\int_{z_0}^\infty dz\frac{dl}{dz}\frac{(1+z_0)^3}{(1+z)^3}\epsilon(\nu,z)\\
  \times\exp{(-\tau_\mathrm{eff}(\nu_0, z_0, z))},
  \label{eqn:flux}
\end{multline}
where
\begin{equation}
  \nu = \nu_0\left(\frac{1+z}{1+z_0}\right).
\end{equation}
Here the effective optical depth is estimated as
\begin{equation}
  \tau_\mathrm{eff}(\nu_0, z_0, z) = \int_{z_0}^z dz^\prime\int_0^\infty dN_\mathrm{HI} f(N_\mathrm{HI}, z^\prime) (1-e^{-\tau_\nu}),
\end{equation}
where $\tau_\nu=\sigma_\nu N_\mathrm{HI}$ and the column density
distribution $f(N_\mathrm{HI}, z)$ is taken to be a power law \citep{2013MNRAS.436.1023B}
\begin{equation}
  f(N_\mathrm{HI}, z) = \frac{A}{N_\mathrm{LL}}\left(\frac{N_\mathrm{HI}}{N_\mathrm{LL}}\right)^{-\beta_N}\left(\frac{1+z}{4.5}\right)^{-\beta_z},
\end{equation}
with $A=0.93$, $\beta_N=1.33$, $\beta_z=1.92$, and
$N_\mathrm{LL}=10^{17.2}$~cm$^{-2}$.  The hydrogen photoionization
rate is then given by
\begin{equation}
  \Gamma_\mathrm{HI}=4\pi\int_{\nu_{912}}^\infty d\nu \frac{j(\nu,z)}{h\nu} \sigma(\nu),
\end{equation}
where $\sigma$ is the photoionization cross-section.  The result is
shown in Figure~\ref{fig:gammapi}, in which as before the solid black
curve shows the hydrogen photoionisate rate in our best-fit global
model, and the yellow curves show a random sample from the global
posterior.  Also, as before, the right panel of
Figure~\ref{fig:gammapi} shows the result when we only account for
quasars brighter than $M_{1450}=-20$, while the left panel shows the
result when the integration limit is $M_{1450}=-18$.  The red points
with error bars in this figure show the measurements of
\citet{2013MNRAS.436.1023B}, derived from the Lyman-$\alpha$ forest,
and the orange points show the measurements by
\citet{2011MNRAS.412.2543C} from quasar proximity zones.  The dashed
green curve shows the hydrogen photoionisation rate evolution in the
model of \citet{2012ApJ...746..125H}, while the solid green curve
shows the contribution of quasars to it.

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{g_18_10oct.pdf}
    \includegraphics[width=0.5\textwidth]{g_20_10oct.pdf}
    \end{tabular}
  \end{center}
  \caption{AGN contribution to the hydrogen photoionisation rate,
    assuming 100\% escape fraction.  Model luminosity functions are
    integrated down to $M_{1450}=-18$ in the left panel and $-20$ in
    the right panel.  The solid black curve shows the hydrogen
    photoionisate rate in our best-fit global model, and the yellow
    curves show a random sample from the global posterior.  The red
    points with error bars in this figure show the measurements of
    \citet{2013MNRAS.436.1023B}, derived from the Lyman-$\alpha$
    forest, and the orange points show the measurements by
    \citet{2011MNRAS.412.2543C} from quasar proximity zones.  The
    dashed green curve shows the hydrogen photoionisation rate
    evolution in the model of \citet{2012ApJ...746..125H}, while the
    solid green curve shows the contribution of quasars to it. }
  \label{fig:gammapi}
\end{figure*}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth,keepaspectratio]{g_puc.pdf}
  \end{center}
  \caption{Photon underproduction crisis.}
\end{figure}

%% \begin{figure*}
%%   \begin{center}
%%     \includegraphics[width=\textwidth,keepaspectratio]{evolution.pdf}
%%   \end{center}
%%   \caption{Parameter evolution from individual fits.  Coloured points
%%     show results when Giallongo quasars are not included.  Black
%%     points show results when Giallongo quasars are included.}
%% \end{figure*}

@@

%% \begin{figure}
%%   \begin{center}
%%     \includegraphics[width=\columnwidth,keepaspectratio]{rhoqso_withGlobal_dense.pdf}
%%   \end{center}
%%   \caption{Alternative presentation of Figure~\ref{fig:rhoqso}.}
%% \end{figure}

%% \begin{figure}
%%   \begin{center}
%%     \includegraphics[width=\columnwidth,keepaspectratio]{rhoqso_diff.pdf}
%%   \end{center}
%%   \caption{Alternative presentation of Figure~\ref{fig:rhoqso}.}
%% \end{figure}

%% \begin{figure}
%%   \begin{center}
%%     \includegraphics[width=\columnwidth,keepaspectratio]{rhoqso_diff_ind.pdf}
%%   \end{center}
%%   \caption{Alternative presentation of Figure~\ref{fig:rhoqso}.}
%% \end{figure}

\section{Hydrogen and helium reionization}

\section{Photon underproduction}

\citet{2014ApJ...789L..32K} reported that the hydrogen photoionization
rate required by simulations to match the observed HI column density
distribution and the effective Ly$\alpha$ opacity at low redshifts
($z\sim 0.1$) is a factor of five (check factor---it looks like about
ten in the plot) higher than that predicted by the model of
\citep{2012ApJ...746..125H}.  This deficit in the hydrogen
photoionization rate at these redshifts, termed `photon
underproduction crisis', has inspired models in which the excess
photonization in explained by blazars (citation) or an evolution in
the density structure of the IGM.  Subsequent simulations by
\citet{2015ApJ...811....3S}, calibrated the observed HI column density
distribution and the effective Ly$\alpha$ optical depth reduced the
severity of the problem by

\section{Conclusion}

We have presented an analysis of the evolution of the UV luminosity
function of AGN from redshift $z=0$ to $7.5$ using a combined
homogenised sample of close to 90,000 UV-selected AGN.  We only
consider quasars with spectroscopic redshifts and reliable
completeness estimates.  These data span magnitudes down to
$M_{1450}=-22$ at high redshifts ($z\gtrsim 3.5$).

\begin{itemize}
\item The AGN luminosity function prefers a double power law
  description at all redshift, except possibly at very high redshifts
  ($z\sim 6$) where only the faint-end of the luminosity function is
  observed.

\item The break magnitude $M_*$ of the double power law luminosity
  function shows a steep brightening towards the high redshift.  The
  break evolves from $M_*<-25$ at $z\sim 2$ to $M_*\sim -29$ at $z\sim
  6$.  Correspondingly, the amplitude $\phi_*$ of the luminosity
  function drops rapidly towards high redshifts from $\phi_*\sim
  10^{-6}$~mag$^{-1}$cMpc$^{-3}$ at $z\sim 2$ to $\phi_*<
  10^{-10}$~mag$^{-1}$cMpc$^{-3}$ at $z\sim 6$.  The break magnitude
  at $z=6$ is very bright, about $-30$, which is why the constraints
  on the bright end slope at these redshifts are weak.

\item The faint end slope $\beta$ of the luminosity function shows
  increased steepening at high redshifts, consistent with the findings
  of \citep{2015AA...578A..83G}.  The faint-end slope is $\beta\sim
  -1.8$ at $z\sim 2$ and steepens to $\beta\sim-2.5$ at $z\sim 6$.
  The bright-end slope $\alpha$ also shows moderate steepening towards
  high redshifts.  However, the constraints on $\alpha$ are weak,
  particularly at the high redshift because of the bright break
  luminosity.

\item A fourteen-parameter model accurately captures the redshift
  evolution of the luminosity function.  In this model, the evolution
  of the break magnitude $M_*$ is described by a cubic polynomial in
  $1+z$, while the evolution of the luminosity function amplitude
  follows a quadratic polynomial in $1+z$.  The evolution of the
  bright-end slope $\alpha$ evolves linearly in $1+z$.  The evolution
  of the faint-end slope $\beta$ is more complicated.  The faint-end
  slope stays roughly constant at lower redshifts $z<3$, steepens
  rapidly at $z\sim 3$--$4$, and then remains constant at higher
  redshifts.

\item An important finding of this work is that there are severe
  systematic errors in observed luminosity function.  We find that the
  most significant systematic bias is at redshift $z\sim 3$ in the
  BOSS sample of quasars.  There is also some evidence of a systematic
  bias in the lowest redshifts $z\lesssim 0.5$.  At $z<2.2$ there is
  some indication of a systematic error in completeness estimates at
  the faintest magnitudes $M_{1450}\sim -19$.

\item In spite of the rapid steepening of the luminosity function at
  high redshifts, our derived LyC emissivity from AGN is quite low.
  This is because of the bright break luminosities, which results in a
  reduced contribution of intermediate brightness quasars to the Lyc
  emissivity as compared to some other determinations in the
  literature.  The LyC emissivity increases with time and peaks at
  $z\sim 2$.  At lower redshifts, the emissivity drops rapidly with
  time.  When our luminoisty function model is extrapolated and
  integrated down to $M_{1450}\sim -20$, the implied LyC emissivity is
  comparable to that in the model of \citet{2007ApJ...654..731H} and
  \citet{2012ApJ...746..125H}.  This indicates that hydrogen
  reionization is unlikely to have caused by AGN alone.

\item We derive photoionization rate by considering the HI column
  density distribution used by \citep{2012ApJ...746..125H}.  We find
  that the HI photoionization rate is significantly lower than that
  required by the Lyman-$\alpha$ forest data at $z>2$.  Our hydrogen
  photoionization rate estimates are also considerably lower (by a
  factor of at least two) than that of \citet{2015AA...578A..83G}.  At
  the lowest redshifts, we find that the hydrogen photoionization rate
  is about a factor of two higher than the estimate of
  \citet{2007ApJ...654..731H} and \citet{2012ApJ...746..125H}.  When
  compared with the recent analysis of \citet{2017MNRAS.467.3172G},
  this photoionisation rate appears to allevaite the photon
  underproduction crisis.  Cite Kollmeier here.

\item Helium reionization in our model occurs at redshift $z\sim
  3$ or $z=3.5$ if the luminosity function is integrated down to
  $M_{1450}=-20$ and $-18$, respectively.

\item Our analysis allows us to analyse the X-ray selected quasars of
  \citet{2015AA...578A..83G} in the larger context of UV-selected
  quasars.  We find that the the faint-end luminosity function
  measurements of \citet{2015AA...578A..83G} at $z=4$--$6$ are
  consitent with the measurements of the luminosity function at
  brighter luminosities.  Stil, our inferred value of the hydrogen
  ionizing luminosity density from quasars at these redshifts is much
  lower that the value derived by \citet{2015AA...578A..83G}.  This
  difference arises due to the difference in the data used in the two
  analyses at intermediate to bright luminosities.  
\end{itemize}

@@

% Add an appendix on comparison with Giallongo.
